
\chapter{Something}

\section{Inverse Method}
Scheme:
\begin{enumerate}
    \item $F_x(x) = \int^x_{-\infty}f_x(x)dx $
    \item $F_x^{-1}(\cdot)$
    \item  $u \sim U[0,1] x~F^{-1}(u)$
\end{enumerate}

\begin{myproof}
    Proof that $X \sim F_x(x)$\\
    \begin{equation}
        \begin{aligned}
            P(X\le x)&= P(F_x^{-1}(u)\le x)&\mbox{}\\[1.25ex]
            P(F_x^{-1}(u)\le x)&=P(u\leF_x(x))&\mbox{}\\[1.25ex]
            P(u\le F_x(x))&=F_x(x)&\mbox{}\\[1.25ex]
        \end{aligned}
    \end{equation}
\end{myproof}

\section{Rejection Method}

Given functions $f(x)$ and  $g(x)$, where  $f(x)$ is more difficult to ???, and there exists $ c > 0$ s.t.  $\forall x \; f(x) \le cg(x)$.\\

\ex{}
{
    Suppose that $f(x) = \frac{e^{-\frac{x^{2}}{2}}}{\sqrt{2\pi}}$ if $g(x)$ is the uniform distribution, the method won't work because f is compactly supported, while g is not. A good choice for  $g(x)$ is the mirrored laplace distribution ... HERE BE EQUATION it is a simple exponential function with absolute x in the exponent.
}
\subsection{Scheme}
\begin{enumerate}
    \item Find $g(x)$ and  $c$ s.t. $f(x) \le cg(x)$ 
    \item Let $x \sim g(x)$ and  $u \sim U[0,1]$ 
    \item Build the pair $(x,cg(x)u)$ 
    \item If $cg(x)u \le f(x)$ take x, otherwise reject it.
\end{enumerate}
\nt
{
    This is a method of creating arbitrary distribution given its probability density function and a well behaved random generator. So if I have a normal number generator, I can use it to generate random numbers along a different distribution for which I have no generator, as long as the above conditions are fulfilled.
}
\clearpage
\ex{Simple example}
{
    HERE BE PLOT OF THE TRIANGLE DISTRIBUTION ON THE INTERVAL [0,1], AND OF A UNIFORM DISTRIBUTION ON THE SAME DOMAIN.
    \begin{equation}
        f(x) = \begin{cases}
            0 \; as x < 0\\
            2x \\
            0 \; as x > 1 
        \end{cases}
    \end{equation}
    \begin{equation}
        g(x) = \begin{cases}
            0 \; as x < 0\\
            1 \\
            0 \; as x > 1 
        \end{cases}
    \end{equation}
 
    HOW DO I DO IT SIDE BY SIDE I FORGOT AAAAAAAAAA\\
    The most appropriate choice for c is $c=2$, as
     \begin{equation}
        \begin{aligned}
            f(x)&\le cg(x)&\mbox{}\\[1.25ex]
            2x  &\le 2\cdot 1&\mbox{True on [0,1]}\\[1.25ex]
        \end{aligned}
    \end{equation}
    Larger values would reject more numbers than necessary.
    ALSO WHY DON'T I HAVE SNIPPETS FOR PLOTTING\\
    The acceptance condition is:
    \begin{equation}
        \begin{aligned}
            cg(x)u&\le f(x)&\mbox{}\\[1.25ex]
            2 \cdot 1 \cdot u&\le 2x&\mbox{}\\[1.25ex]
            u&\le x&\mbox{}\\[1.25ex]
            
        \end{aligned}
    \end{equation}
    \nt
    {
        In this example, x is just the uniform distribution, but it can be any distribution at all. On the other hand, by the definition of the rejection method u is always uniform.
    }

    \nt
    {
        $N \sim \sin \cos{u}$ ???? what
    }
}
